### 2025-07-18 actuarial.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - chapter overviews

@@

>>>>

This chapter compares [judgment](#sec-humanJudgment) versus [actuarial approaches](#sec-actuarialPrediction) to prediction in fantasy football (and other domains).

@@

::: {.content-visible unless-format="html:js"}

@TPS2017 provides a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions.

### 2025-07-18 cognitive-bias.qmd
20250717 - chapter overviews

@@

In conclusion, there are many [heuristics](#sec-heuristics), [cognitive bias](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
It is prudent to be aware of these common biases and to work to counteract them.
For instance, look for information that challenges or disconfirms your beliefs, and work to incorporate this information into your beliefs.
Do your best to pursue observation, data analysis, and data interpretation with an open mind.
You never know what important information you might discover if you go in with an open mind.
Pay attention to [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation) when considering whether a player's performance may [regress to the mean](#sec-fallaciesRegression).
If the strong performances appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
There is considerable luck in fantasy football.
Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.

>>>>

This chapter provides an overview of many [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions in fantasy football.

@@

In conclusion, there are many [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
It is prudent to be aware of these common biases and to work to counteract them.
For instance, look for information that challenges or disconfirms your beliefs, and work to incorporate this information into your beliefs.
Do your best to pursue observation, data analysis, and data interpretation with an open mind.
You never know what important information you might discover if you go in with an open mind.
Pay attention to [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation) when considering whether a player's performance may [regress to the mean](#sec-fallaciesRegression).
If the strong performances appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
There is considerable luck in fantasy football.
Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.

### 2025-07-18 evaluating-prediction-accuracy.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - chapter overviews

@@

>>>>

::: {.content-visible unless-format="html:js"}

> "*Nothing ruins fantasy more than reality.*" – [Renee Miller, Ph.D.](https://www.youtube.com/watch?v=gmpLFWs5ae0&t=1430) [@Yahoo2024]

This chapter provides an overview of ways to evaluate the accuracy of predictions.
In addition, we evaluate the accuracy of fantasy football projections.

@@

In terms of the accuracy of historical fantasy football projections, projections showed moderate accuracy in terms of [discrimination](#sec-discrimination) in predicting players' fantasy performance.
However, projections were not highly accurate.
Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
Projection sources appeared to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity, which is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
Crowd-averaged projections tend to be more accurate than individual sources of projections.

### 2025-07-18 mythbusters.qmd
20250708 - track chunk execution times
20250717 - chapter overviews

@@

```{r}
library("petersenlab")
library("nflreadr")
library("lme4")
library("lmerTest")
library("MuMIn")
library("emmeans")
library("tidyverse")
```

@@

summary(mixedModel_qbr)
MuMIn::r.squaredGLMM(mixedModel_qbr)
emmeans::emmeans(mixedModel_qbr, "contractYear")

@@

summary(mixedModelAge_qbr)
MuMIn::r.squaredGLMM(mixedModelAge_qbr)
emmeans::emmeans(mixedModelAge_qbr, "contractYear")
```

@@

summary(mixedModel_ptsAdded)
MuMIn::r.squaredGLMM(mixedModel_ptsAdded)
emmeans::emmeans(mixedModel_ptsAdded, "contractYear")

@@

summary(mixedModelAge_ptsAdded)
MuMIn::r.squaredGLMM(mixedModelAge_ptsAdded)
emmeans::emmeans(mixedModelAge_ptsAdded, "contractYear")
```

@@

summary(mixedModel_epaPass)
MuMIn::r.squaredGLMM(mixedModel_epaPass)
emmeans::emmeans(mixedModel_epaPass, "contractYear")

@@

summary(mixedModelAge_epaPass)
MuMIn::r.squaredGLMM(mixedModelAge_epaPass)
emmeans::emmeans(mixedModelAge_epaPass, "contractYear")
```

@@

summary(mixedModel_fantasyPtsPass)
MuMIn::r.squaredGLMM(mixedModel_fantasyPtsPass)
emmeans::emmeans(mixedModel_fantasyPtsPass, "contractYear")

@@

summary(mixedModelAge_fantasyPtsPass)
MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsPass)
emmeans::emmeans(mixedModelAge_fantasyPtsPass, "contractYear")
```

@@

summary(mixedModel_ypc)
MuMIn::r.squaredGLMM(mixedModel_ypc)
emmeans::emmeans(mixedModel_ypc, "contractYear")

@@

summary(mixedModelAge_ypc)
MuMIn::r.squaredGLMM(mixedModelAge_ypc)
emmeans::emmeans(mixedModelAge_ypc, "contractYear")
```

@@

summary(mixedModel_epaRush)
MuMIn::r.squaredGLMM(mixedModel_epaRush)
emmeans::emmeans(mixedModel_epaRush, "contractYear")

@@

summary(mixedModelAge_epaRush)
MuMIn::r.squaredGLMM(mixedModelAge_epaRush)
emmeans::emmeans(mixedModelAge_epaRush, "contractYear")
```

@@

summary(mixedModel_fantasyPtsRush)
MuMIn::r.squaredGLMM(mixedModel_fantasyPtsRush)
emmeans::emmeans(mixedModel_fantasyPtsRush, "contractYear")

@@

summary(mixedModelAge_fantasyPtsRush)
MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsRush)
emmeans::emmeans(mixedModelAge_fantasyPtsRush, "contractYear")
```

@@

summary(mixedModel_receivingYards)
MuMIn::r.squaredGLMM(mixedModel_receivingYards)
emmeans::emmeans(mixedModel_receivingYards, "contractYear")

@@

summary(mixedModelAge_receivingYards)
MuMIn::r.squaredGLMM(mixedModelAge_receivingYards)
emmeans::emmeans(mixedModelAge_receivingYards, "contractYear")
```

@@

summary(mixedModel_epaReceiving)
MuMIn::r.squaredGLMM(mixedModel_epaReceiving)
emmeans::emmeans(mixedModel_epaReceiving, "contractYear")

@@

summary(mixedModelAge_epaReceiving)
MuMIn::r.squaredGLMM(mixedModelAge_epaReceiving)
emmeans::emmeans(mixedModelAge_epaReceiving, "contractYear")
```

@@

summary(mixedModel_fantasyPtsReceiving)
MuMIn::r.squaredGLMM(mixedModel_fantasyPtsReceiving)
emmeans::emmeans(mixedModel_fantasyPtsReceiving, "contractYear")

@@

summary(mixedModelAge_fantasyPtsReceiving)
MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsReceiving)
emmeans::emmeans(mixedModelAge_fantasyPtsReceiving, "contractYear")
```

@@

>>>>

In this chapter, we put a popular fantasy football belief to the test.
We evaluate whether the widely held belief is true that players perform better during a contract year.

@@

```{r}
library("petersenlab")
library("nflreadr")
library("lme4")
library("lmerTest")
library("performance")
library("emmeans")
library("tidyverse")
```

@@

summary(mixedModel_qbr)
performance::r2(mixedModel_qbr)
emmeans::emmeans(mixedModel_qbr, "contractYear")

@@

summary(mixedModelAge_qbr)
performance::r2(mixedModelAge_qbr)
emmeans::emmeans(mixedModelAge_qbr, "contractYear")
```

@@

summary(mixedModel_ptsAdded)
performance::r2(mixedModel_ptsAdded)
emmeans::emmeans(mixedModel_ptsAdded, "contractYear")

@@

summary(mixedModelAge_ptsAdded)
performance::r2(mixedModelAge_ptsAdded)
emmeans::emmeans(mixedModelAge_ptsAdded, "contractYear")
```

@@

summary(mixedModel_epaPass)
performance::r2(mixedModel_epaPass)
emmeans::emmeans(mixedModel_epaPass, "contractYear")

@@

summary(mixedModelAge_epaPass)
performance::r2(mixedModelAge_epaPass)
emmeans::emmeans(mixedModelAge_epaPass, "contractYear")
```

@@

summary(mixedModel_fantasyPtsPass)
performance::r2(mixedModel_fantasyPtsPass)
emmeans::emmeans(mixedModel_fantasyPtsPass, "contractYear")

@@

summary(mixedModelAge_fantasyPtsPass)
performance::r2(mixedModelAge_fantasyPtsPass)
emmeans::emmeans(mixedModelAge_fantasyPtsPass, "contractYear")
```

@@

summary(mixedModel_ypc)
performance::r2(mixedModel_ypc)
emmeans::emmeans(mixedModel_ypc, "contractYear")

@@

summary(mixedModelAge_ypc)
performance::r2(mixedModelAge_ypc)
emmeans::emmeans(mixedModelAge_ypc, "contractYear")
```

@@

summary(mixedModel_epaRush)
performance::r2(mixedModel_epaRush)
emmeans::emmeans(mixedModel_epaRush, "contractYear")

@@

summary(mixedModelAge_epaRush)
performance::r2(mixedModelAge_epaRush)
emmeans::emmeans(mixedModelAge_epaRush, "contractYear")
```

@@

summary(mixedModel_fantasyPtsRush)
performance::r2(mixedModel_fantasyPtsRush)
emmeans::emmeans(mixedModel_fantasyPtsRush, "contractYear")

@@

summary(mixedModelAge_fantasyPtsRush)
performance::r2(mixedModelAge_fantasyPtsRush)
emmeans::emmeans(mixedModelAge_fantasyPtsRush, "contractYear")
```

@@

summary(mixedModel_receivingYards)
performance::r2(mixedModel_receivingYards)
emmeans::emmeans(mixedModel_receivingYards, "contractYear")

@@

summary(mixedModelAge_receivingYards)
performance::r2(mixedModelAge_receivingYards)
emmeans::emmeans(mixedModelAge_receivingYards, "contractYear")
```

@@

summary(mixedModel_epaReceiving)
performance::r2(mixedModel_epaReceiving)
emmeans::emmeans(mixedModel_epaReceiving, "contractYear")

@@

summary(mixedModelAge_epaReceiving)
performance::r2(mixedModelAge_epaReceiving)
emmeans::emmeans(mixedModelAge_epaReceiving, "contractYear")
```

@@

summary(mixedModel_fantasyPtsReceiving)
performance::r2(mixedModel_fantasyPtsReceiving)
emmeans::emmeans(mixedModel_fantasyPtsReceiving, "contractYear")

@@

summary(mixedModelAge_fantasyPtsReceiving)
performance::r2(mixedModelAge_fantasyPtsReceiving)
emmeans::emmeans(mixedModelAge_fantasyPtsReceiving, "contractYear")
```

@@

```{r}
player_statsContractsQBRBWRTE_seasonal <- player_statsContracts_seasonal %>% 
  dplyr::filter(
    position_group %in% c("QB","RB","WR","TE"),
    games >= 5, # keep only player-season combinations in which QBs played at least 5 games
    season >= 2011) # keep only seasons since 2011 (when most contract data are available)
```

```{r}
mixedModel_fantasyPts <- lmerTest::lmer(
  fantasyPoints ~ contractYear + position_group + (1 | player_id),
  data = player_statsContractsQBRBWRTE_seasonal,
  control = lmerControl(optimizer = "bobyqa")
)

summary(mixedModel_fantasyPts)
performance::r2(mixedModel_fantasyPts)
emmeans::emmeans(mixedModel_fantasyPts, "contractYear")

mixedModelAge_fantasyPts <- lmerTest::lmer(
  fantasyPoints ~ contractYear + position_group + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
  data = player_statsContractsQBRBWRTE_seasonal,
  control = lmerControl(optimizer = "bobyqa")
)

summary(mixedModelAge_fantasyPts)
performance::r2(mixedModelAge_fantasyPts)
emmeans::emmeans(mixedModelAge_fantasyPts, "contractYear")
```
